---
title: "Intro to fishflux"
author: "Nina M. Schiettekatte"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: xelatex
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8, fig.height=6
)
```
## Introduction
  
The `fishflux` package provides a tool to model fluxes of C (carbon), N (nitrogen) and P (phosphorus) in fish. It combines basic principles from elemental stoichiometry and metabolic theory. The package offers a user friendly interface to make nutrient dynamic modelling available for anyone. `fishflux` is mostly targeted towards fish ecologists, wishing to predict nutrient ingestion, egestion and excretion to study fluxes of nutrients and energy. 

Main assets:
  
* Provides functions to model fluxes of Carbon, Nitrogen and Phosphorus for fish with or without the MCMC sampler provided by stan.
* Provides some tools to find the right parameters as inputs into the model
* Provides a plotting function to illustrate results

## Installing and loading fishflux

`fishflux` uses Markov Chain Monte Carlo simulations provided by [stan](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started). Therefore, the first step is to install [stan](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).

### GitHub

The best way to install the latest development version of `fishflux` is to install it from GitHub. 

``` r
install.packages("devtools")
devtools::install_github("nschiett/fishflux", dependencies=TRUE)
library(fishflux)
```
### CRAN

`fishflux` will be available on CRAN in the future:
  
``` r
install.packages("fishflux")
library(fishflux)
```

### Downloaded package file

Another option is to download the source file available on GitHub [here](https://github.com/nschiett/fishflux).

``` r
install.packages(path_to_fishflux_file, repos = NULL, type="source")
library(fishflux)
```

## How to use fishflux?

`fishflux` is designed to follow three simple steps:

* Find the right input parameters
* Run the model simulation with those input parameters
* Plot the model results

### Find parameters

Before running the model, the parameters have to be specified. Below, there is a table showing all parameters needed to run the model simulation. `fishflux` provides several functions to find some of these parameters, but note that others have to be provided by the user at this stage. Ideally, all parameters should also have a standard deviation, so that uncertainty of parameters can be reflected in the model output parameters. 
\newline

```{r,message=FALSE,echo=FALSE,results='asis'}
knitr::kable(data.frame(
  Parameters = c("C","N","P","Fc","Fn","Fp","AEc","AEn","AEp","Linf","k","t0","lwa","lwb","w_prop","Tn","Tp","B0","a","f","asp","troph","temp"), 
  Description = c("Carbon content of fish (%)", "Nitrogen content of fish (%)","Phosphorus content of fish (%)","C content of food (%)","N content of food (%)","P content of food (%)", "C-specific absorption efficiency", "N-specific absorption efficienct","P-specific absorption efficiency", "Von Bertalanffy growth parameter","Von Bertalanffy growth parameter","Von Bertalanffy growth parameter", "Parameter length-weight relationship","Parameter length-weight relationship","Ratio between dry weight and wet weight of fish","N-specific turnover rate","P-specific turnover rate","Normalisation constant for resting metabolic rate", "Resting metabolic rate mass-scaling exponent","Activity scope","Aspect ratio","Trophic level","Temperature (°C)")
),caption = "Overview all input variables")
```

\newline

A good place to start is checking if you are using the correct scientific name of your fish of interest. The function `name_errors` will tell you if the species name is correct. This function can be useful, especially when working with larger databases.  

\newline
```{r,message=TRUE,echo=TRUE}
# example
fishflux::name_errors("Zebrazoma scopas")
```
\newline
Once the species names are verified and/or corrected we can continue with specifying some parameters. 

The `find_lw` function searches fishbase to find length-weight relationship parameters `lw_a` and `lw_b` extracted from @Froese2013. 
\newline
```{r,message=TRUE,echo=TRUE}
# example
fishflux::find_lw("Zebrasoma scopas")
```
\newline
The model uses Von Beralanfy patameters to approximate growth. A quick way to get available information from fishbase is the function `growth_params()`. This can be a good indication, but users should interpret these estimates with a critical eye, as they come from disparate sources of varying accuracy. Alternatively, it is advised to use otolith readings. The function `oto_growth()` allows for direct estimation of growth parameters from otolith data, using a hierarchical bayesian model and stan. In the case of absence of otolith data, one might consider extracting standardised estimations from Morais & Bellwood (2018).
\newline
```{r,message=FALSE,echo=TRUE}
# example
# The option otolith=TRUE filters out sources that used otoliths for the estimation of growth parameters
fishflux::growth_params("Sargocentron microstoma", otolith = FALSE)

# Load example otolith data
sarmi <- fishflux::sarmi
# Fit bayesian model
oto_growth <- fishflux::oto_growth(length = sarmi$length, age = sarmi$age, id = sarmi$id, linf_m = 16, k_m = 0.4, iter = 2000, warmup = 1000)
oto_growth$result

## Note that in this example, there is a large difference between fishbase estimates and otolith analysis. This illustrates how important it is to check the quality of your parameter estimates.
```
\newline

For metabolic parameters B0 and a, there is a function to extract estimations from Barneche & Allen 2018 Ecology Letters doi: 10.1111/ele.12947. These parameters are for the best model (Model 2 in the paper online supplementary material) of fish resting metabolic rates reported in the paper, which also includes trophic level as a covariate. Please cite Barneche & Allen (2018), when using these values. 
\newline
```{r,echo=TRUE,,cache=TRUE,results='hide',message=FALSE}
# example
metabolism <- fishflux::metabolism("Zebrasoma scopas", troph_m = 2, temp = 27)
```
```{r}
print(metabolism)
```

\newline
Further, there are a couple more basic functions to get an indication of parameters that are available on fishbase such as `trophic_level()` and `aspect_ratio()`.
\newline
Note that it is always better to get the approximations through analysis, measurements and otolith analysis over parameters extracted from functions, such as `growth_params()`, `trophic_level()` and `aspect_ratio()`. 
\newline
To get an overview of all parameters, `fishflux` provides a wrapper function `model_parameters`.
\newline
```{r,cache=TRUE,results='hide',message=FALSE}
# example
zebsco <- fishflux::model_parameters("Zebrasoma scopas", family = "Acanthuridae", temp = 27)
```
```{r}
print(zebsco)
```
\newline
All other parameters have to be provided by the user. For more information on how to aquire these paramers, take a look at ("this paper")add reference to methods paper).


### Run model

Once all the parameters are collected, we can run the model through `cnp_model_mcmc()`. Note that this model can be run with or without specifying the sd's of each parameter. If a sd of a certain parameter is not provided, it will be automatically set to a very low value (1⁻10). As mentioned before, it is advisable to include uncertainty of parameters. Fishflux is designed to use the mcmc sampler in order to predict uncertainty of predictions. Alternatively, to get a quick prediction of averages, without depending on stan, the function `cnp_model()` can be used. 
\newline
```{r, message=FALSE}
## load the example parameters for Zebrasoma scopas, a list
param_zebsco <- fishflux::param_zebsco
## Run the model, specifying the target length(s) and the parameter list
model <- fishflux::cnp_model_mcmc(TL = 5:20, param = param_zebsco)
model$lim
```
\newline
The object 'model' now contains all the samples generated from the mcmc simulation and a summary of all parameters generated. `model$lim` shows the limiting element at each length. To extract certain parameters of interest, use the `extract()` function. Predictions for fluxes of C, N and P are all in g/day.
\newline
```{r}
fishflux::extract(model, c("N_ex","P_ex"))
```
\newline
### Plot results 

To visualize the main outputs of the model, `fishflux` contains a plotting function. It is also possible to specify the output parameter or element of interest. The lines show the mean of the predicted parameters and the segments illustrate the 95% CI.
\newline
```{r, message=FALSE, warning=FALSE}
## General overview:
fishflux::plot_cnp(model)
## plot certain parameters with 'option':
fishflux::plot_cnp(model, option = "egestion")
## plot one element with 'display':
fishflux::plot_cnp(model, option = c("excretion", "egestion", "growth"), display = "p")
```
\newline
  
### Sensitivity
  
The function `sensitivity()` looks at how the distribution of the input variables affect the uncertainty of the model predictions. 
Basically, the model is run for each input parameter, while keeping all the others fixed. The output of the function gives a matrix of sd's for all model predictions (columns), depending on theinput variables (rows). 
The input parameters and output variables of interest can be specified by arguments "par" and "out" respectively.
\newline 
```{r, message=FALSE}
## General overview:
fishflux::sensitivity(TL = 10, param = param_zebsco, par = c("Fn_sd", "Fp_sd", "C_sd", "N_sd", "P_sd", "k_sd"), out = c("N_ex", "P_ex", "C_in", "C_g", "N_in", "P_in"))
```
\newline
## More information

For more information on the theoretical framework of the model, read paper.
Every function of `fishflux` has a help page with more documentation.
In the case of errors, bugs or discomfort, you are invited to raise an issue at GitHub (should add link). 
`fishflux` is always in development and we are happy to take your comments or suggestions into consideration. 
